The quench action approach for integrable models: A Monte Carlo implementation

### 2015-10-02 draft/fs_qa_v3.tex
rev-02/10/2015

Here we review some Bethe ansatz results for the spin-$\frac{1}{2}$ Heisenberg 
($XXX$) chain. Specifically, in section~\ref{sec:1.1} we introduce the model. 
Its eigenstates (Bethe states) and the Bethe equations are discussed in 
section~\ref{sec:1.2}. Section~\ref{sec:1.3} focuses on the string hypothesis 
and the so-called Bethe-Gaudin-Takahashi (BGT) equations. The form of the 
BGT equations in the thermodynamic limit is discussed in section~\ref{sec:1.4}. 
Finally, in section~\ref{sec:1.5} we provide some exact formulas for the 
local conserved charges of the model.

@@

In the Bethe ansatz framework~\cite{bethe-1931,taka-book} the generic eigenstate 
of~\eref{xxx-ham} (Bethe state) in the sector with $M$ particles can be written as 
%
\begin{equation}
\label{ba-eig}
|\Psi_M\rangle=\sum\limits_{1\le x_1<x_2<\dots<x_M\le L}A_M(x_1,x_2,
\dots,x_M)|x_1,x_2,\dots,x_M\rangle,
\end{equation}
%
where the sum is over the positions $\{x_i\}_{i=1}^M$ of the particles, and $A_M(x_1,
x_2,\dots,x_M)$ is the eigenstate amplitude corresponding to the particles 
being at positions $x_1,x_2,\dots, x_M$. Here $A_M(x_1,x_2,\dots, x_M)$ is 
given as 
%
\begin{equation}
\label{ba_amp}
A_M(x_1,x_2,\dots,x_M)\equiv\sum\limits_{\sigma\in S_M}\exp\Big[i
\sum\limits_{j=1}^Mk_{\sigma_j}x_j+i\sum\limits_{i<j}\theta_{\sigma_i,\sigma_j}
\Big], 
\end{equation}
%
where the outermost summation is over the permutations $S_M$ of the so-called 
quasi-momenta $\{k_\alpha\}_{\alpha=1}^M$. The two-particle scattering phases 
$\theta_{\alpha,\beta}$ are defined as 
%
\begin{equation}
\label{s_phases}
\theta_{\alpha,\beta}\equiv \frac{1}{2i}\log\Big[-\frac{e^{ik_\alpha+ik_\beta}-
2e^{ik_\alpha}+1}{e^{ik_\alpha+ik_\beta}-2e^{ik_\beta}+1}\Big].
\end{equation}
%
The eigenenergy associated to the eigenstate~\eref{ba-eig} is  
%
\begin{equation}
\label{ba-ener}
E=\sum\limits_{\alpha=1}^M(\cos(k_\alpha)-1). 
\end{equation}
%
The quasi-momenta $k_\alpha$ are obtained by solving the so-called Bethe 
equations~\cite{bethe-1931}
%
\begin{equation}
\label{ba-eq}
e^{ik_\alpha L}=\prod\limits^M_{\beta\ne\alpha}\Big[-\frac{1-2e^{
ik_\alpha}-e^{ik_\alpha+ik_\beta}}{1-2e^{ik_\beta}-e^{ik_\alpha+
ik_\beta}}\Big].
\end{equation}
%
It is useful to  introduce the rapidities $\{\lambda_\alpha\}_{\alpha=1}^M$ as 
%
\begin{equation}
\label{rap}
k_\alpha=\pi-2\arctan(\lambda_\alpha)\quad\mbox{mod}\, 2\pi.
\end{equation}
%
Taking the logarithm on both sides in~\eref{ba-eq}, and using~\eref{rap}, 
one obtains the Bethe equations in logarithmic form as 
%
\begin{equation}
\label{ba-eq-log}
\arctan(\lambda_\alpha)=\frac{\pi}{L}J_\alpha+\frac{1}{L}\sum\limits_{
\beta\ne\alpha}\arctan\Big(\frac{\lambda_\alpha-\lambda_\beta}{2}\Big),
\end{equation}
%
where $-L/2<J_\alpha\le L/2$ are the so-called Bethe quantum numbers. The 
$J_\alpha$ are half-integers and integers for $L-M$ even and odd, 
respectively.

Finally, one should remark that $M$-particle eigenstates corresponding to 
{\it finite} rapidities are eigenstates with maximum allowed magnetization 
(highest-weight eigenstates) $S_T^z=L/2-M=S_T$, with $S_T$ the total spin. 
Due to the $SU(2)$ invariance of~\eref{xxx-ham}, all the states in the same 
$S_T$ multiplet and different $-S_T\le S_T^z\le S_T$ are eigenstates of the 
$XXX$ chain with the same energy eigenvalue. These eigenstates (descendants) 
are obtained by multiple applications of the total-spin lowering operator 
$S_T^-\equiv\sum_iS_i^-$ on the highest-weight eigenstates. In the Bethe ansatz 
framework the rapidities of a generic $M$-particle descendant eigenstate with 
$S_T^T=L/2-M'$, with $M'<M$, are obtained by supplementing the $M$ rapidities 
of the highest-weight state with $M'-M$  infinite rapidities. We anticipate 
that descendant eigenstates have non-zero overlap with the zero-momentum 
N\'eel state (cf. section~\ref{sec:2}).

@@

The string centers $\lambda_{n;\gamma}$ are obtained by solving the so-called 
Bethe-Gaudin-Takahashi equations
%
\begin{equation}
\label{bgt-eq}
2L\theta_n(\lambda_{n;\gamma})=2\pi I_{n;\gamma}+\sum\limits_{(m,
\beta)\ne(n,\gamma)}\Theta_{m,n}(\lambda_{n;\gamma}-\lambda_{m;\beta}).  
\end{equation}
%
The generalized scattering phases $\Theta_{m,n}$ read 
%
\begin{eqnarray}
\label{Theta}
\nonumber\fl\Theta_{m,n}(x)\equiv\left\{\begin{array}{cc}
\theta_{|n-m|}(x)+\!\!\!\!\!\sum
\limits_{r=1}^{(n+m-|n-m|-1)/2}\!\!\!\!\!2\theta_{|n-m|+2r}(x)
+\theta_{n+m}(x) & \quad\mbox{if}
\quad n\ne m\\\fl\sum\limits_{r=1}^{n-1}2\theta_{2r}(x)+
\theta_{2n}(x) & \quad\mbox{if}\quad n=m
\end{array}\right.
\end{eqnarray}
%
with $\theta_\alpha(x)\equiv 2\arctan(x/\alpha)$. Here $I_{n;\gamma}$ are the 
Bethe-Takahashi quantum numbers associated with $\lambda_{n;\gamma}$. 
The solutions of~\eref{bgt-eq}, and the Bethe states thereof, are naturally 
classified according to their ``string content'' ${\mathcal S}\equiv\{s_n
\}_{n=1}^M$, with $s_n$ the number of $n$-strings. Clearly, the constraint 
$\sum_{n=1}^{M}n s_n=M$ has to be satisfied. It can be shown that 
the BGT quantum numbers $I_{n;\gamma}$ associated with the $n$-strings are 
integers and half-integers for $L-s_n$ odd and even, respectively. 
An upper bound for the BGT quantum numbers can be derived as~\cite{taka-book} 
%
\begin{equation}
|I_{n;\gamma}|\le I^{(MAX)}_{n}\equiv\frac{1}{2}(L-1-\sum
\limits_{m=1}^Mt_{m,n}s_m),
\label{bt-qn-bound}
\end{equation}
%
where $t_{m,n}\equiv 2\mbox{min}(n,m)-\delta_{m,n}$. Using the string hypothesis 
Bethe states energy eigenvalue~\eref{ba-ener} becomes
%
\begin{equation}
\label{ener-str}
E=-\sum_{n,\gamma}\frac{2n}{\lambda_{n;\gamma}^2+n^2}. 
\end{equation}
%

@@

In the thermodynamic limit $L\to\infty$ at fixed finite particle density $M/L$ 
the solutions of the BGT equations~\eref{bgt-eq} become dense. One then defines 
the BGT root distributions for the $n$-strings as $\pmb{\rho}\equiv\{\rho_n(
\lambda)\}_{n=1}^\infty$, with $\rho_n(\lambda)\equiv\lim_{L\to\infty}[
\lambda_{n;\gamma+1}-\lambda_{n;\gamma}]^{-1}$. Thus the  BGT 
equations~\eref{bgt-eq} become an infinite set of coupled non-linear integral 
equations for the $\rho_n(\lambda)$ as 
%
\begin{equation}
\label{bgt-th}
a_n(\lambda)=\rho_n(\lambda)+\rho^h_n(\lambda)+\sum_m(T_{n,m}*\rho_m)
(\lambda),
\end{equation}
%
where $\rho_n^{h}$ are the hole-distributions, and the functions $a_n(\lambda)$ 
are defined as 
%
\begin{equation}
a_n(x)\equiv\frac{1}{\pi}\frac{n}{x^2+n^2}. 
\end{equation}
%
In~\eref{bgt-th} $T_{n,m}*\rho_m$ denotes the convolution 
%
\begin{equation}
(T_{n,m}*\rho_m)(\lambda)\equiv\int_{-\infty}^{+\infty}T_{n,m}(\lambda-\lambda')
\rho_{m}(\lambda'),
\end{equation}
%
with the matrix $T_{n,m}(x)\equiv\Theta'(x)$ being dfined as 
%
\begin{eqnarray}
\nonumber\fl T_{m,n}(x)\equiv\left\{\begin{array}{cc}
a_{|n-m|}(x)+\!\!\!\!\!\sum
\limits_{r=1}^{(n+m-|n-m|-1)/2}\!\!\!\!\!2a_{|n-m|+2r}(x)
+a_{n+m}(x) & \quad\mbox{if}
\quad n\ne m\\\fl\sum\limits_{r=1}^{n-1}2a_{2r}(x)+
a_{2n}(x) & \quad\mbox{if}\quad n=m
\end{array}\right.
\end{eqnarray}
%
For a generic smooth enough observable ${\mathcal O}$ in the thermodynamic 
limtit its expectation value is expected to become a functional of the root 
densities $\pmb{\rho}$ as $\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}\rangle$.

Moreover, for all the local observables considered in this work the contribution of the 
different type of strings factorize and one can write 
%
\begin{equation}
\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}\rangle=\sum_{n=1}^\infty
\int_{-\infty}^{+\infty}d\lambda \rho_n(\lambda) {\mathcal O}_n(\lambda), 
\end{equation}
% 
with ${\mathcal O}_n(\lambda)$ the ... .

@@

The functional integral in~\eref{qa-d-ensemble} can be evaluated in the limit 
$L\to\infty$ using the saddle point approximation. One has to minimize the 
functional ${\mathcal F}(\pmb{\rho})$ defined as 
%
\begin{equation}
L{\mathcal F}(\pmb{\rho})\equiv 2\Re\log\langle\pmb{\rho}|\Psi_0
\rangle+S_{YY}(\pmb{\rho}(\lambda))  
\end{equation}
% 
with respect to $\pmb{\rho}$, i.e., solving $\delta{\mathcal F}(\pmb{\rho})/\delta
\pmb{\rho}|_{\pmb{\rho}=\pmb{\rho^*}}=0$, under the constraint that the thermodynamic BGT 
equations~\eref{bgt-th} hold. Finally, one obtains from~\eref{qa-d-ensemble} that in the 
thermodynamic limit 
%
\begin{equation}
\label{obs-th}
\langle{\mathcal O}\rangle=\langle\pmb{\rho^*}|{\mathcal O}|\pmb{\rho^*}\rangle. 
\end{equation}
%
Remarkably, for the quench with initial state the N\'eel state $|\Psi_0\rangle=|N\rangle$ the 
saddle point root distributions ${\rho_n^*(\lambda)}_{n=1}^\infty$ can be obtained 
analytically~\cite{brockmann-2014}. The first few are given as 
%
\begin{eqnarray}
\label{rho1-sp}
\fl\quad\rho^*_1(\lambda)=\frac{8(4+\lambda^2)}{\pi(19+3\lambda^2)(1+6\lambda^2+
\lambda^4)}\\\label{rho2-sp}
\fl\quad\rho_2^*(\lambda)=\frac{8\lambda^2(9+\lambda^2)(4+3\lambda^2)}{\pi(2+\lambda^2)
(16+14\lambda^2+\lambda^4)(256+132\lambda^2+9\lambda^4)}\\\label{rho3-sp}
\fl\quad\rho_3^*(\lambda)=\frac{8(1+\lambda^2)^2(5+\lambda^2)(16+\lambda^2)(21+\lambda^2)}
{\pi(19+3\lambda^2)(9+624\lambda^2+262\lambda^4+32\lambda^6+\lambda^8)
(509+5\lambda^2(26+\lambda^2))}.
\end{eqnarray}
%

@@

In this section we discuss generic features of the overlaps between the 
eigenstates (Bethe states) of the Heisenberg spin chain and the N\'eel state 
(cf.~\eref{neel}). We exploit the Bethe ansatz solution of the chain (see 
section~\ref{sec:1}) as well as exact results for the N\'eel overlaps (see 
section~\ref{sec:2}). We focus on finite chains with $L\lesssim 40$ sites, 
considering, for any $L$, the full Hilbert space of the chain. 
The only Bethe states having, in principle, non zero N\'eel overlap are the 
so-called parity-invariant Bethe states (see section~\ref{sec:2}). We 
denote their total number as $Z_{Neel}$. Crucially, here we restrict 
ourselves to the subset of parity-invariant Bethe states. These correspond 
to solutions of the Bethe-Gaudin-Takahashi (BGT) equations~\eref{bgt-eq} 
that do not contain zero-momentum strings. We denote the total number of 
these eigenstates as $\widetilde Z_{Neel}$. Both $Z_{Neel}$ and $\widetilde 
Z_{Neel}$ are given in terms of the chain length $L$ by simple combinatorial 
formulas that we provide.

@@

Interestingly, the fraction of eigenstates with no zero-momentum strings, i.e.,  
$\widetilde Z_{Neel}/Z_{Neel}$, is vanishing as $L^{-1/2}$ in the thermodynamic 
limit, meaning that zero-momentum strings eigenstates are dominant, at least 
in number, for large chains. This has dramatic consequences, for instance, at 
the level of the overlap sum rules for the local conservation laws of the model. 
Precisely, although the sum rules are fixed by the initial (N\'eel) state 
expectation value of the local charges, we observe striking violations for 
any finite chain. Moreover, all the sum rules exhibit vanishing behavior as 
$L^{-1/2}$ upon increasing the chain size, reflecting the same vanishing behavior 
as $\widetilde Z_{Neel}/Z_{Neel}$. A similar scenario holds for the overlaps with 
the Majumdar-Ghosh state, where, however, excluding the zero-momentum strings leads 
to a $1/L$ behavior. This demonstrates that for finite chains the physical effects 
of zero-momentum strings eigenstates are not negligible.

@@

Here we overview the Bethe ansatz results for the N\'eel overlaps with the 
eigenstates of the $XXX$ chain. The total number of partiy-invariant 
eigenstates $Z_{Neel}$ having, in principle, non-zero N\'eel overlap is given 
as 
%
\begin{equation}
\label{zneel1}
Z_{Neel}=2^{\frac{L}{2}-1}+\frac{1}{2}B\Big(\frac{L}{2},\frac{L}{4}\Big)+1, 
\end{equation}
%
with $B(n,m)\equiv n!/(m!(n-m)!)$ the Newton binomial. The proof of~\eref{zneel1} 
is obtained by counting all the parity-invariant BGT quantum number configurations, 
and it is reported in~\ref{app-1}. Note that $Z_{Neel}$ provides only an upper 
bound for the number of eigenstates with non-zero N\'eel overlap, as it is clear 
from the exact diagonalization results shown in Table~\ref{table:neel}. This is 
because parity-invariant eigenstates with a single zero-momentum even-length 
string, which are included in~\eref{zneel1}, have identically zero N\'eel 
overlap~\cite{brockmann-2014}. This is not related to the symmetries of the N\'eel 
state, but to the ``accidental'' vanishing of the prefactor in the overlap 
formula~\eref{neel-ov}. Finally, after excluding the zero-momentum strings 
eigenstates, the total number of remaining eigenstates $\widetilde Z_{Neel}$, 
which are the ones considered here, is given as (see~\ref{app-2} for the proof) 
%
\begin{equation}
\label{ztilde}
\widetilde Z_{Neel}=B\Big(\frac{L}{2},\frac{L}{4}\Big).
\end{equation}
%
An overview of generic features of the overlaps is given in Figure~\ref{fig0:neel-ov} 
(a) plotting the squared N\'eel overlaps $|\langle\lambda|N\rangle|^2$ with the $XXX$ 
chain eigenstates $|\lambda\rangle$ versus the energy density $E/L$. The figure shows 
results for chains with $26\le L\le 38$ sites. The data are obtained by generating all 
the relevant parity-invariant BGT quantum numbers, and solving the associated BGT 
equations~\eref{bgt-eq}, to obtain the rapidities of $XXX$ chain eigenstates. The 
N\'eel overlaps are calculated numerically using~\eref{neel-ov}. Note that for $L=38$ 
from~\eref{ztilde} the total number of overlap shown in the Figure is $\widetilde Z_{Neel}
\sim 10^5$.

>>>>

Here we review some Bethe ansatz results for the spin-$\frac{1}{2}$ Heisenberg 
($XXX$) chain. Specifically, in section~\ref{sec:1.1} we introduce the model. 
Its eigenstates (Bethe states) and the Bethe equations are discussed in 
section~\ref{sec:1.2}. Section~\ref{sec:1.3} focuses on the string hypothesis 
and the so-called Bethe-Gaudin-Takahashi (BGT) equations. The form of the 
BGT equations in the thermodynamic limit is discussed in section~\ref{sec:1.4}. 
Finally, in section~\ref{sec:1.5} we provide the exact formulas for the 
local conserved charges of the model.

@@

In the Bethe ansatz framework~\cite{bethe-1931,taka-book} the generic eigenstate 
of~\eref{xxx-ham} (Bethe state) in the sector with $M$ particles can be written as 
%
\begin{equation}
\label{ba-eig}
|\Psi_M\rangle=\sum\limits_{1\le x_1<x_2<\dots<x_M\le L}A_M(x_1,x_2,
\dots,x_M)|x_1,x_2,\dots,x_M\rangle,
\end{equation}
%
where the sum is over the positions $\{x_i\}_{i=1}^M$ of the particles, and $A_M(x_1,
x_2,\dots,x_M)$ is the eigenstate amplitude corresponding to the particles 
being at positions $x_1,x_2,\dots, x_M$. Here $A_M(x_1,x_2,\dots, x_M)$ is 
given as 
%
\begin{equation}
\label{ba_amp}
A_M(x_1,x_2,\dots,x_M)\equiv\sum\limits_{\sigma\in S_M}\exp\Big[i
\sum\limits_{j=1}^Mk_{\sigma_j}x_j+i\sum\limits_{i<j}\theta_{\sigma_i,\sigma_j}
\Big], 
\end{equation}
%
where the outermost summation is over the permutations $S_M$ of the so-called 
quasi-momenta $\{k_\alpha\}_{\alpha=1}^M$. The two-particle scattering phases 
$\theta_{\alpha,\beta}$ are defined as 
%
\begin{equation}
\label{s_phases}
\theta_{\alpha,\beta}\equiv \frac{1}{2i}\log\Big[-\frac{e^{ik_\alpha+ik_\beta}-
2e^{ik_\alpha}+1}{e^{ik_\alpha+ik_\beta}-2e^{ik_\beta}+1}\Big].
\end{equation}
%
The eigenenergy associated with the eigenstate~\eref{ba-eig} is  
%
\begin{equation}
\label{ba-ener}
E=\sum\limits_{\alpha=1}^M(\cos(k_\alpha)-1). 
\end{equation}
%
The quasi-momenta $k_\alpha$ are obtained by solving the so-called Bethe 
equations~\cite{bethe-1931}
%
\begin{equation}
\label{ba-eq}
e^{ik_\alpha L}=\prod\limits^M_{\beta\ne\alpha}\Big[-\frac{1-2e^{
ik_\alpha}-e^{ik_\alpha+ik_\beta}}{1-2e^{ik_\beta}-e^{ik_\alpha+
ik_\beta}}\Big].
\end{equation}
%
It is useful to  introduce the rapidities $\{\lambda_\alpha\}_{\alpha=1}^M$ as 
%
\begin{equation}
\label{rap}
k_\alpha=\pi-2\arctan(\lambda_\alpha)\quad\mbox{mod}\, 2\pi.
\end{equation}
%
Taking the logarithm on both sides in~\eref{ba-eq}, and using~\eref{rap}, 
one obtains the Bethe equations in logarithmic form as 
%
\begin{equation}
\label{ba-eq-log}
\arctan(\lambda_\alpha)=\frac{\pi}{L}J_\alpha+\frac{1}{L}\sum\limits_{
\beta\ne\alpha}\arctan\Big(\frac{\lambda_\alpha-\lambda_\beta}{2}\Big),
\end{equation}
%
where $-L/2<J_\alpha\le L/2$ are the so-called Bethe quantum numbers. The 
$J_\alpha$ are half-integers and integers for $L-M$ even and odd, 
respectively~\cite{taka-book}.

Importantly, the $M$-particle Bethe states~\eref{ba-eig} corresponding to 
{\it finite} rapidities are eigenstates with maximum allowed magnetization 
(highest-weight eigenstates) $S_T^z=L/2-M=S_T$, with $S_T$ the total spin. 
Due to the $SU(2)$ invariance of~\eref{xxx-ham}, all the states in the same 
$S_T$ multiplet and different $-S_T\le S_T^z\le S_T$ are eigenstates of the 
$XXX$ chain with the same energy eigenvalue. These eigenstates (descendants) 
are obtained by multiple applications of the total-spin lowering operator 
$S_T^-\equiv\sum_iS_i^-$ on the highest-weight states. In the Bethe ansatz 
framework, given a highest-weight eigenstate with $M'$ particles (i.e., $M'$ 
finite rapidities), its descendants are obtained by supplementing the $M'$ 
rapidities with infinite rapidities. We anticipate that descendant eigenstates 
are important here since they have non-zero overlap with the zero-momentum 
N\'eel state (cf. section~\ref{sec:2}).

@@

The string centers $\lambda_{n;\gamma}$ are obtained by solving the so-called 
Bethe-Gaudin-Takahashi equations~\cite{taka-book}
%
\begin{equation}
\label{bgt-eq}
2L\theta_n(\lambda_{n;\gamma})=2\pi I_{n;\gamma}+\sum\limits_{(m,
\beta)\ne(n,\gamma)}\Theta_{m,n}(\lambda_{n;\gamma}-\lambda_{m;\beta}).  
\end{equation}
%
Here the generalized scattering phases $\Theta_{m,n}$ read 
%
\begin{eqnarray}
\label{Theta}
\nonumber\fl\Theta_{m,n}(x)\equiv\left\{\begin{array}{cc}
\theta_{|n-m|}(x)+\!\!\!\!\!\sum
\limits_{r=1}^{(n+m-|n-m|-1)/2}\!\!\!\!\!2\theta_{|n-m|+2r}(x)
+\theta_{n+m}(x) & \quad\mbox{if}
\quad n\ne m\\\fl\sum\limits_{r=1}^{n-1}2\theta_{2r}(x)+
\theta_{2n}(x) & \quad\mbox{if}\quad n=m
\end{array}\right.
\end{eqnarray}
%
with $\theta_\alpha(x)\equiv 2\arctan(x/\alpha)$, and $I_{n;\gamma}$  the 
Bethe-Takahashi quantum numbers associated with $\lambda_{n;\gamma}$. 
The solutions of~\eref{bgt-eq}, and the Bethe states~\eref{ba-eig} thereof, 
are naturally classified according to their ``string content'' ${\mathcal S}
\equiv\{s_n\}_{n=1}^M$, with $s_n$ the number of $n$-strings. Clearly, the 
constraint $\sum_{n=1}^{M}n s_n=M$ has to be satisfied. It can be shown that 
the BGT quantum numbers $I_{n;\gamma}$ associated with the $n$-strings are 
integers and half-integers for $L-s_n$ odd and even, respectively. 
Moreover, an upper bound for the BGT quantum numbers can be 
derived as~\cite{taka-book} 
%
\begin{equation}
|I_{n;\gamma}|\le I^{(MAX)}_{n}\equiv\frac{1}{2}(L-1-\sum
\limits_{m=1}^Mt_{m,n}s_m),
\label{bt-qn-bound}
\end{equation}
%
where $t_{m,n}\equiv 2\mbox{min}(n,m)-\delta_{m,n}$. Using the string hypothesis 
Bethe states energy eigenvalue~\eref{ba-ener} becomes
%
\begin{equation}
\label{ener-str}
E=-\sum_{n,\gamma}\frac{2n}{\lambda_{n;\gamma}^2+n^2}. 
\end{equation}
%

@@

In the thermodynamic limit $L\to\infty$ at fixed finite particle density $M/L$ 
the solutions of the BGT equations~\eref{bgt-eq} become dense. One then defines 
the BGT root distributions for the $n$-strings as $\pmb{\rho}\equiv\{\rho_n(
\lambda)\}_{n=1}^\infty$, with $\rho_n(\lambda)\equiv\lim_{L\to\infty}[
\lambda_{n;\gamma+1}-\lambda_{n;\gamma}]^{-1}$. As a consequence, the  BGT 
equations~\eref{bgt-eq} become an infinite set of coupled non-linear integral 
equations for the $\rho_n(\lambda)$ as 
%
\begin{equation}
\label{bgt-th}
a_n(\lambda)=\rho_n(\lambda)+\rho^h_n(\lambda)+\sum_m(T_{n,m}*\rho_m)
(\lambda),
\end{equation}
%
where $\rho_n^{h}$ are the so-called hole-distributions, and the functions 
$a_n(\lambda)$ are defined as 
%
\begin{equation}
a_n(x)\equiv\frac{1}{\pi}\frac{n}{x^2+n^2}. 
\end{equation}
%
In~\eref{bgt-th} $T_{n,m}*\rho_m$ denotes the convolution 
%
\begin{equation}
(T_{n,m}*\rho_m)(\lambda)\equiv\int_{-\infty}^{+\infty}T_{n,m}(\lambda-\lambda')
\rho_{m}(\lambda'),
\end{equation}
%
with the matrix $T_{n,m}(x)\equiv\Theta'(x)$ being dfined as 
%
\begin{eqnarray}
\nonumber\fl T_{m,n}(x)\equiv\left\{\begin{array}{cc}
a_{|n-m|}(x)+\!\!\!\!\!\sum
\limits_{r=1}^{(n+m-|n-m|-1)/2}\!\!\!\!\!2a_{|n-m|+2r}(x)
+a_{n+m}(x) & \quad\mbox{if}
\quad n\ne m\\\fl\sum\limits_{r=1}^{n-1}2a_{2r}(x)+
a_{2n}(x) & \quad\mbox{if}\quad n=m
\end{array}\right.
\end{eqnarray}
%
Given a generic, smooth enough, observable ${\mathcal O}$, in thermodynamic 
limit its eigenstate expectation value is replaced by a functional of the root 
densities $\pmb{\rho}$ as $\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}\rangle$. 
Moreover, we should stress that for all the local observables considered here 
the contribution of the different type of strings factorize, and $\langle\pmb{
\rho}|{\mathcal O}|\pmb{\rho}\rangle$ becomes 
%
\begin{equation}
\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}\rangle=\sum_{n=1}^\infty
\int_{-\infty}^{+\infty}d\lambda \rho_n(\lambda) {\mathcal O}_n(\lambda), 
\end{equation}
% 
with ${\mathcal O}_n(\lambda)$ the .

@@

The functional integral in~\eref{qa-d-ensemble} can be evaluated in the limit 
$L\to\infty$ using the saddle point approximation. One has to minimize the 
functional ${\mathcal F}(\pmb{\rho})$ defined as 
%
\begin{equation}
\label{qa-term}
L{\mathcal F}(\pmb{\rho})\equiv 2\Re\log\langle\pmb{\rho}|\Psi_0
\rangle+S_{YY}(\pmb{\rho}(\lambda))  
\end{equation}
% 
with respect to $\pmb{\rho}$, i.e., solving $\delta{\mathcal F}(\pmb{\rho})/\delta
\pmb{\rho}|_{\pmb{\rho}=\pmb{\rho^*}}=0$, under the constraint that the thermodynamic BGT 
equations~\eref{bgt-th} hold. Finally, one obtains from~\eref{qa-d-ensemble} that in the 
thermodynamic limit 
%
\begin{equation}
\label{obs-th}
\langle{\mathcal O}\rangle=\langle\pmb{\rho^*}|{\mathcal O}|\pmb{\rho^*}\rangle. 
\end{equation}
%
Remarkably, for the quench with initial state the N\'eel state $|\Psi_0\rangle=|N\rangle$ the 
saddle point root distributions ${\rho_n^*(\lambda)}_{n=1}^\infty$ can be obtained 
analytically~\cite{brockmann-2014}. The first few are given as 
%
\begin{eqnarray}
\label{rho1-sp}
\fl\quad\rho^*_1(\lambda)=\frac{8(4+\lambda^2)}{\pi(19+3\lambda^2)(1+6\lambda^2+
\lambda^4)}\\\label{rho2-sp}
\fl\quad\rho_2^*(\lambda)=\frac{8\lambda^2(9+\lambda^2)(4+3\lambda^2)}{\pi(2+\lambda^2)
(16+14\lambda^2+\lambda^4)(256+132\lambda^2+9\lambda^4)}\\\label{rho3-sp}
\fl\quad\rho_3^*(\lambda)=\frac{8(1+\lambda^2)^2(5+\lambda^2)(16+\lambda^2)(21+\lambda^2)}
{\pi(19+3\lambda^2)(9+624\lambda^2+262\lambda^4+32\lambda^6+\lambda^8)
(509+5\lambda^2(26+\lambda^2))}.
\end{eqnarray}
%

@@

In this section we discuss generic features of the overlaps between the 
eigenstates (Bethe states) of the Heisenberg spin chain and the N\'eel state 
(cf.~\eref{neel}). We exploit the Bethe ansatz solution of the chain (see 
section~\ref{sec:1}) as well as exact results for the N\'eel overlaps (see 
section~\ref{sec:2}). We focus on finite chains with $L\lesssim 40$ sites, 
considering, for any $L$. The only Bethe states having, in principle, non 
zero N\'eel overlap are the so-called parity-invariant Bethe states (see 
section~\ref{sec:2}). We denote their total number as $Z_{Neel}$. Crucially, 
here we restrict ourselves to the parity-invariant Bethe states that do not 
contain zero-momentum strings. We denote the total number of these eigenstates 
as $\widetilde Z_{Neel}$. Both $Z_{Neel}$ and $\widetilde Z_{Neel}$ are given 
in terms of the chain length $L$ by simple combinatorial formulas that we 
provide.

Interestingly, the fraction of eigenstates with no zero-momentum strings, i.e.,  
$\widetilde Z_{Neel}/Z_{Neel}$, is vanishing as $L^{-1/2}$ in the thermodynamic 
limit, meaning that zero-momentum strings eigenstates are dominant in number 
for large chains. This, however, has dramatic consequences, for instance, at the 
level of the overlap sum rules for the local conservation laws of the model. 
Precisely, although the sum rules are fixed by the initial (N\'eel) state 
expectation value of the local charges, we observe striking violations for 
any finite chain. Moreover, all the sum rules exhibit vanishing behavior as 
$L^{-1/2}$ upon increasing the chain size, reflecting the same vanishing behavior 
as $\widetilde Z_{Neel}/Z_{Neel}$. A similar scenario holds for the overlaps with 
the Majumdar-Ghosh state, where excluding the zero-momentum strings leads to a $1/L$ 
behavior. This demonstrates that for finite chains the physical effects of 
zero-momentum strings eigenstates are not negligible.

@@

@@

Here we overview the Bethe ansatz results for the N\'eel overlaps with the 
eigenstates of the $XXX$ chain. The total number of partiy-invariant 
eigenstates $Z_{Neel}$ having, in principle, non-zero N\'eel overlap is given 
as 
%
\begin{equation}
\label{zneel1}
Z_{Neel}=2^{\frac{L}{2}-1}+\frac{1}{2}B\Big(\frac{L}{2},\frac{L}{4}\Big)+1, 
\end{equation}
%
with $B(n,m)\equiv n!/(m!(n-m)!)$ the Newton binomial. The proof of~\eref{zneel1} 
is obtained by counting all the parity-invariant BGT quantum number configurations, 
and it is reported in~\ref{app-1}. Note that $Z_{Neel}$ provides only an upper 
bound for the number of eigenstates with non-zero N\'eel overlap, as it is clear 
from the exact diagonalization results shown in Table~\ref{table:neel}. This is 
because parity-invariant eigenstates with a single zero-momentum even-length 
string, which are included in~\eref{zneel1}, have identically zero N\'eel 
overlap~\cite{brockmann-2014}. This is not related to the symmetries of the N\'eel 
state, but to an ``accidental'' vanishing of the prefactor in the overlap 
formula~\eref{neel-ov}. Finally, after excluding the zero-momentum strings 
eigenstates, the total number of remaining eigenstates $\widetilde Z_{Neel}$, 
which are the ones considered here, is given as (see~\ref{app-1.2} for the proof) 
%
\begin{equation}
\label{ztilde}
\widetilde Z_{Neel}=B\Big(\frac{L}{2},\frac{L}{4}\Big).
\end{equation}
%
An overview of generic features of the overlaps is given in Figure~\ref{fig0:neel-ov} 
(a) plotting the squared N\'eel overlaps $|\langle\lambda|N\rangle|^2$ with the $XXX$ 
chain eigenstates $|\lambda\rangle$ versus the energy density $E/L$. The figure shows 
results for chains with $26\le L\le 38$ sites. The data are obtained by generating all 
the relevant parity-invariant BGT quantum numbers, and solving the associated BGT 
equations~\eref{bgt-eq}, to obtain the rapidities of $XXX$ chain eigenstates. Finally, 
the overlaps are calculated numerically using~\eref{neel-ov}. Note that for $L=38$ 
from~\eref{ztilde} the total number of overlap shown in the Figure is $\widetilde Z_{Neel}
\sim 10^5$.
